---
title: "p8105_hw3_zz3167"
author: "Zhiyi Zhu"
date: "2023-10-08"
output: github_document
---

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```


```{r}
knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = .6,
  out.width = "90%",
  message = FALSE, 
  warning = FALSE
)
```

## Problem 1

### Load the data

```{r}
library(p8105.datasets)
data("instacart")
```

### Write a short description of the dataset

* The dataset contains `r nrow(instacart)` observations of `r length(unique(instacart$order_id))` unique users with `r ncol(instacart)` variables, which form a total of `r nrow(instacart) * ncol(instacart)` data points.
* The dataset consists of information about products ordered through Instacart. Each row corresponds to a product and contains details such as the product's ID (`product_id`), the order's ID (`order_id`), the order in which the item was added to the cart (`add_to_cart_order`), whether the item has been ordered before (`reordered`), and the customer's ID (`user_id`).
*  Each product is described by its name (`product_name`), the aisle it belongs to (`aisle`), the aisle's ID (`aisle_id`), and the department it falls under (`department`). For instance, the product's name could be something like "Bulgarian Yogurt," it belongs to the "yogurt" aisle with an aisle ID(120), and its department is "dairy eggs." Similarly, another product might be a "Organic Celery Hearts" in the "fresh vegetables" aisle with aisle ID(83), belonging to the "produce" department.

### How many aisles are there, and which aisles are the most items ordered from?

```{r}
aisles_num = 
  select(instacart,aisle_id) |>
  unique() |>
  nrow()

aisles_desc_order_df = 
  instacart |>
  group_by(aisle) |>
  summarize(n_obs = n()) |>
  arrange(desc(n_obs)) 
  
aisles_max =
  aisles_desc_order_df |>
  head(1) |>
  pull(aisle)
```

There's total `r aisles_num` aisles, among which `r aisles_max` are the most items ordered from.

### Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
aisles_desc_order_df |>
  janitor::clean_names() |>
  filter(n_obs > 10000) |>
  ggplot(aes(x = reorder(aisle, n_obs), y = n_obs)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Number of Items Ordered at Each Aisle",
    x = "Aisle",
    y = "Number of items ordered",
    caption = "Note: limited to aisles with more than 10000 items ordered"
  )
```

#### Comment on the results:

* The plot indicates that the largest number of items ordered comes from the "fresh vegetables" and "fresh fruits" aisles. 

### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart |>
  janitor::clean_names() |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

#### Comment on the results:

* In the "Packaged Vegetables Fruits" aisle, "Organic Baby Spinach" is the most popular product with a count of 9784. "Organic Raspberries" and "Organic Blueberries" are the 2nd and 3rd most popular in the same aisle, with counts of 5546 and 4966 respectively.
* In the "Baking Ingredients" aisle, "Light Brown Sugar" is the top product with a count of 499. "Pure Baking Soda" and "Cane Sugar" are the 2nd and 3rd most popular baking ingredients, with counts of 387 and 336 respectively.
* In the "Dog Food Care" aisle, 'Snack Sticks Chicken & Rice Recipe Dog Treats' is the most popular with a count of 30. 'Organix Chicken & Brown Rice Recipe' and 'Small Dog Biscuits' are the 2nd and 3rd most popular dog food care products, with counts of 28 and 26 respectively.

### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers 

```{r}
instacart |>
  janitor::clean_names() |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(order_dow, product_name) |>
  summarise(mean_hour_of_the_day = mean(order_hour_of_day)) |>
  mutate(
    order_dow = case_match(
      order_dow,
      0 ~ 'Sunday',
      1 ~ 'Monday',
      2 ~ 'Tuesday',
      3 ~ 'Wednesday',
      4 ~ 'Thursday',
      5 ~ 'Friday',
      6 ~ 'Saturday'
    )) |>
  pivot_wider(names_from = order_dow, values_from = mean_hour_of_the_day) |>
  knitr::kable(digits = 2)
```

#### Comment on the results:

* In general, customers tend to order 'Coffee Ice Cream' later in the day, especially on Tuesday, Wednesday, and Thursday. For 'Pink Lady Apples', the orders tend to be placed earlier in the day, Usually around noon.

## Problem 2

### Load the data

```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

### Do some data cleaning

```{r}
brfss_new_df = 
  brfss_smart2010 |>
  janitor::clean_names() |>
  rename(state = locationabbr, location = locationdesc) |>
  filter(
    topic == "Overall Health",
    response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) |>
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"), order = TRUE))
```

### Answer the following questions and comment on the results.

#### In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
states_2002 = 
  brfss_new_df |>
  filter(year == "2002") |>
  group_by(state) |>
  summarise(n_location = n_distinct(location)) |>
  filter(n_location >= 7) |>
  pull(state)

states_2010 = 
  brfss_new_df |>
  filter(year == "2010") |>
  group_by(state) |>
  summarise(n_location = n_distinct(location)) |>
  filter(n_location >= 7) |>
  pull(state)
```

#### Comment on the results:

In 2002, `r states_2002` were observed at 7 or more locations.

In 2010, `r states_2010` were observed at 7 or more locations. 

#### Construct a dataset that meets the criteria and make a plot.

```{r}
brfss_excellent = 
  brfss_new_df |>
  filter(response == "Excellent") |>
  group_by(year, state) |>
  summarise(mean_data_value = mean(data_value, na.rm = TRUE))

brfss_excellent |>
  ggplot(aes(x = year, y = mean_data_value, group = state, color = state)) +
  geom_line() +
  labs(
    title ="Excellent Responses Variation Across Different States",
    x = "Year",
    y = "Average Value") +
  theme_bw()
```

#### Comment on the results:

* From the spaghetti plot we can see that the mean value varies between `r min(brfss_excellent$mean_data_value)` and `r max(brfss_excellent$mean_data_value)`. For some states, the fluctuation is considerable.

#### Make a two-panel plot 

```{r}
brfss_new_df |>
  filter(year %in% c(2006,2010),state == "NY") |>
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  facet_wrap(.~ year) +
  labs(title = "Distribution of Data Value by Response in NY State for 2006 and 2010",
       x = "Response",
       y = "Data Value")
```

#### Comment on the results:

* Comparing the two boxplots, we can find that they have similar distribution. In 2006 and 2010, "very good" response corresponded to the highest average data value, and "poor" response corresponded to the lowest average data value.


## Problem 3

### Load, tidy, merge and organize the data sets

```{r}
nhanes_accel = 
  read_csv("data/nhanes_accel.csv") |>
  janitor::clean_names()

nhanes_covar =
  read_csv("data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names()

nhanes_df = 
  full_join(nhanes_covar, nhanes_accel, by = "seqn") |>
  filter(age >= 21) |>
  drop_na(sex, age, bmi, education) |>
  mutate(sex = case_match(sex, 1 ~ "male", 2 ~ "female"),
         sex = factor(sex, levels = c("male", "female"), ordered = TRUE),
         education = case_match(education,
                                1 ~ "Less than high school",
                                2 ~ "High school equivalent",
                                3 ~ "More than high school"),
         education = factor(education, levels = c("Less than high school", "High school equivalent", "More than high school"), ordered = TRUE))
```

### Produce a reader-friendly table

```{r}
nhanes_df |>
  group_by(sex, education) |>
  summarise(n_obs = n()) |>
  pivot_wider(
    names_from = education,
    values_from = n_obs
  ) |>
  knitr::kable()
```

#### Comment on these items

From the table, we can find that the number of people with the educational category of "More than high school" is the largest and the number of people with the educational category of "Less than high school" is the least. Across all education categories, men outnumbered women at "High school equivalent" and "More than high school" categories. At "Less than high school" category, women outnumber men slightly. 

### Create a visualization of the age distributions

```{r}
nhanes_df |>
  ggplot(aes(x = sex, y = age, fill = sex)) +
  geom_boxplot() +
  labs(title = "Age Distributions",
       x = "Sex",
       y = "Age") +
  facet_grid(.~education)
```

#### Comment on these items

<mark>Comment


### Create a total activity variable for each participant

```{r}
total_activity_df = 
  nhanes_df|>
  pivot_longer(
    cols = starts_with("min"),
               names_to = "time",
               values_to = "min_act")|>
  group_by(seqn)|>
  summarize(total_activity=sum(min_act)) 

combine_activity_df = 
  full_join(nhanes_df, total_activity_df, by ="seqn") |>
  select(-starts_with("min"))
```

#### Comment on these items

<mark>Comment

### Plot these total activities (y-axis) against age (x-axis)

```{r}
combine_activity_df |>
  ggplot(aes(x = age, y = total_activity, color = sex)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_grid(.~education) +
  labs(title = "Total Activity against by Age, Sex, and Education Level",
       x = "Age",
       y = "Total Activity")
```

#### Comment on these items

<mark>Comment

### Make a three-panel plot that shows the 24-hour activity time courses

```{r}
nhanes_df |>
  pivot_longer(
    cols = starts_with("min"),
    names_to = "time",
    values_to = "min_act"
  ) |>
  ggplot(aes(x = time, y = min_act, color = sex)) +
  geom_point(alpha = .5) +
  geom_smooth() +
  facet_grid(.~education) +
  labs(
    title = "24-Hour Activity Time Courses",
    x = "time",
    y = "activity",
    caption = "Data from nhanes dataset"
  )
```





